---
title: "Untitled"
author: "Rocio Meza"
date: "April 11, 2017"
output: ioslides_presentation
font: smaller
---

<style type="text/css"> body, td { font-size: 5px; } code.r{ font-size: 10px; } pre { font-size: 5px } </style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=FALSE}
library(readr)
vf <- read_csv("~/Documents/Statistics/voterfile.csv")
Regression.5 <- glm(vh12g ~ vh12p + vh10g + vh10p + 
                      vh08g +	vh08p + vh06g + vh06p + vh04g + 	
                      vh04p + vh00g + age + factor(party), 
                    family = binomial(link = "logit"), data = vf)

vf$vote_prob <- predict(Regression.5, vf, type="response")
#vf$vote_prob

#Creating Column that show '1' or '0' for Will Vote or Will NOT Vote (as numeric/not character)
vf$vote <- ifelse(vf$vote_prob >= 0.5,"1","0") 
vf$vote <- as.numeric(vf$vote)

mytable.party <- table(vf$party, vf$vote) # A is row, B is columns 
mytable.party

margin.table(mytable.party, 1) # A frequencies (summed over B) 
```

## R Markdown
{ font-size: 5px; }
This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Prediction 
Using a Generalized Linear Model and a sample of 5,000 registered votes, I was able to predict the voter turnout (projected to vote or not) for the 2014 General Election in Nevada. 

In my sample, my model predicted that appoximately *27%* of registered voters will vote in the 2014 General Election, while approximately *73%* will not. Below we can see the percentages of those will vote (1) for each political party. 
```{r chunk_name, echo=FALSE}
prop.table(mytable.party, 1) # row percentages 
```

## Slide with Plot

```{r pressure}
plot(pressure)
```

