---
title: "0ptimus Voter Turnout"
author: "Rocio Meza"
date: "April 11, 2017"
output: pdf_document
---

```{r}
#Loading Data + Naming File 
library(readr)
vf <- read_csv("~/Documents/Statistics/voterfile.csv")
```


# Logistic Regression (GLM) : Using Historical Voting Data 

I chose to use the 2012 General Election data (vh12g) as the dependent variable in my logistic regression and made the past voting history as the indepdent variables. I chose the latest General Election in my data because past historical data is one of the *best* predictors of voter turnout. In addition, I chose the binomial family of error models along with its default identity (loggit). 

*Assumptions Made for Logistic Regression*: 

* Dependent variable is binary. 

* Since logistic regression assumes that P(Y=1) is the probability of the event occurring, it is necessary that the dependent variable is coded accordingly.  That is, for a binary regression, the factor level 1 of the dependent variable should represent the desired outcome, which is correct in this case! 

* The error terms need to be independent.

* Logistic regression assumes linearity of independent variables and log odds.

* Sample size has to be big enough. ML needs at least 10 cases per independent variable, some statisticians recommend at least 30 cases for each parameter to be estimated.

In the results we see that 'vh02g' and 'vh02p'were not statistically significant.

**It is important to note that I am using a considering anything significant as long as it's below the significane level of 0.05**
```{r}
# First Regression With ALL of Voting History before 2012 General Election 
# Using 2012 General Election to predict 2014 General Election 

Regression.1 <- glm(vf$vh12g ~ vf$vh12p + vf$vh10g + vf$vh10p + 
                    vf$vh08g +	vf$vh08p + vf$vh06g + vf$vh06p + vf$vh04g + 	
                    vf$vh04p + vf$vh02g + vf$vh02p + vf$vh00g + vf$vh00p, 
                    family = binomial(link = "logit"))


summary(Regression.1)
```

Next, I removed 'vh02g' and 'vh02p.' 

In this example, we have a value of 57710 on 49999 degrees of freedom. Including the independent variables, decreased the deviance to 25861 points on 49988 degrees of freedom, a significant reduction in deviance.

The Residual Deviance was reduced by **31849** with a loss of 11 degrees of freedom.
```{r}
#Remove 'vh02g' and 'vh02p'
Regression.2 <- glm(vf$vh12g ~ vf$vh12p + vf$vh10g + vf$vh10p + 
                      vf$vh08g +	vf$vh08p + vf$vh06g + vf$vh06p + vf$vh04g + 	
                      vf$vh04p + vf$vh00g + vf$vh00p, 
                    family = binomial(link = "logit"))
summary(Regression.2)
```

Then, I considered adding the variables 'age' and 'party' into the glm. 

Here we see that age IS significant to the model. However, Natural Law and Other parties 
are NOT.

In this example, we have a value of 57687 on 49982 degrees of freedom. Including the independent variables, decreased the deviance to 25599 points on 49963 degrees of freedom, a significant reduction in deviance.

The Residual Deviance has reduced by **32088** with a loss of 19 degrees of freedom.

```{r}
#Including Age & Party into the Mix. 
Regression.3 <- glm(vh12g ~ vh12p + vh10g + vh10p + 
                      vh08g +	vh08p + vh06g + vh06p + vh04g + 	
                      vh04p + vh00g + vh00p + age + factor(party), 
                    family = binomial(link = "logit"), data = vf)

summary(Regression.3)
```

Considering that some political parties had no statistical significance to the model, I decided to exclude 'party' to see if it would create a more fitted model. 

'vh00p' was not proven to be significant in this model either. 

For this example, we have a value of 57687 on 49982 degrees of freedom. Including the independent variables, decreased the deviance to 25624 points on 49970 degrees of freedom, a significant reduction in deviance.

The Residual Deviance has reduced by **32063** with a loss of 12 degrees of freedom. It was not as great as the loss of the Residual Deviance (**32088**) of the previous model that INCLUDED 'party.' 
```{R}
#Including Age ONLY into the mix.  
Regression.4 <- glm(vh12g ~ vh12p + vh10g + vh10p + 
                      vh08g +	vh08p + vh06g + vh06p + vh04g + 	
                      vh04p + vh00g + vh00p + age, 
                    family = binomial(link = "logit"), data = vf)


summary(Regression.4)
```

At the end, I chose to create a model that would exclude vh00p, 'vh02g', and 'vh02p.' It included the rest of the voting histories, age, and party. 

For this example, we have a value of 57687 on 49982 degrees of freedom. Including the independent variables, decreased the deviance to 25599 points on 49964 degrees of freedom, a significant reduction in deviance.

The Residual Deviance has reduced by **32088** with a loss of 18 degrees of freedom. This model has the highest reduction of residual deviance. 
```{r}
#Our Choice - 
Regression.5 <- glm(vh12g ~ vh12p + vh10g + vh10p + 
                      vh08g +	vh08p + vh06g + vh06p + vh04g + 	
                      vh04p + vh00g + age + factor(party), 
                    family = binomial(link = "logit"), data = vf)
summary(Regression.5)
```

# Code for Predicting Turnout for Each Indvidual 
```{r}
#Creating Column with Predictions 
vf$vote_prob <- predict(Regression.5, vf, type="response")
#vf$vote_prob

#Creating Column that show '1' or '0' for Will Vote or Will NOT Vote (as numeric/not character)
vf$vote <- ifelse(vf$vote_prob >= 0.5,"1","0") 
vf$vote <- as.numeric(vf$vote)
```

*Exporting to CSV file*
```{r}
write.csv(vf, file = "~/Documents/Statistics/vf.csv", row.names = FALSE)
```

#Interpreting Results 

**Considering that 0ptimus frequently works with Republicans and some Democrats, I have decided to focus on these two parties, but also to include information about other parties as well.**

First I examine my 'vote' column, which has either an '0' or '1'. It turns out that it has NA's, which is okay in this case, because table() ignores NA's. Then I calculate the proportion of Voters and Non-Voters. In this sample of 49983 total registered voters, my model predicted that appoximately *27%* of registered voters will vote in the 2014 General Election, while approximately *73%* will not. 
```{r}
table(vf$vote14)
#Checking for NA because the table does not equal to 5,000
any(is.na(vf$vote))

Total <- 49983

Proportion.Voters <- 13262/ Total 
Proportion.Voters
Proportion.NonVoters <- 36721/ Total 
Proportion.NonVoters
```

I simply made frequency/margin/proportion tables to explore the number of voters who will cast their ballot for the 2014 General Election for each party. 

As we can see, if it was only democrats vs. republicans, the republican candidates would get more votes. However, we see a large number of registered non-partisan people who plan to vote (1255). 

If the model goes as predicted, the democrats would need to gain the support of all the non-partisan voters who are predicted to vote and votes from other political parties. In other words, the democrats would need more than 1279 votes to win.  

**This is a hypothetical. There are 1,424,417 registered voters in Nevada. Therefore, I have included the project number of votes in my slide using statistics from Office of Nevada Secretary of State Ross Miller - May 2014 Voter Registration Statistics.**
```{r}
#Looking at Voters/Non-Votes for Party 
mytable.party <- table(vf$party, vf$vote) # A is row, B is columns 
mytable.party

margin.table(mytable.party, 1) # A frequencies (summed over B) 

prop.table(mytable.party, 1) # row percentages 
# Within each party, republicans have the highest predicted turnout. 
```

I decided to explore Age vs. Party. I created a boxplot for a visual representation of the data. Those in quartiles 2 & 3 for Democrats were around 37 - 65 years old, while Republicans had voters around 42 - 67. 
```{r}
#Age vs. Party Boxplot 
library(ggplot2)
p <- ggplot(vf, aes(party, age))
p + geom_boxplot() + coord_flip()
```

I created bar plots for dwelling type, marital status, networth, ethnicity, homeowner status, and congressional district. I filled in the bars with party affiliations. I also accompanied the plots with frequency/margin/proportion tables.

Results Summary: 

* Single family dwellings are primarily made up of republicans and democrats. 

* Republican and Democratic voters are mainly found in the $100000-249999 and $250000-499999 income brackets. 

* Most registered voters are European. The second largest ethnicity is Hispanic. In the Hispanic group, the majority are Democrats: 
+ 53% are Dem
+ 22% are Non-Partisan 
+ 20% are Rep

* The majority of married registered voters are Republicans 


* Both Democrats and Republicans are almost equally likely to be homeowners.

* District 2 has the highest number of Republicans. District 4 has the highest number of Democrats. 
+ District 1 has more Dems (50%)
+ District 2 has more Rep (32%)
+ District 3 has more Dem (37%) & about (36%) Rep 
+ District 4 has more Dem (53%)

```{r}
#Dwelling Type vs. Party
qplot(factor(dwellingtype), data=vf, geom="bar", fill=factor(party))
mytable.dwelling<- table(vf$party, vf$dwellingtype) # A is row, B is columns 
mytable.dwelling

margin.table(mytable.dwelling, 2) # A frequencies (summed over B) 

prop.table(mytable.dwelling, 1) # row percentages 
```

```{r}
# NetWorth vs. Party 
qplot(factor(net_worth), data=vf, geom="bar", fill=factor(party))
mytable.networth<- table(vf$party, vf$net_worth) # A is row, B is columns 
mytable.networth

margin.table(mytable.networth, 1) # A frequencies (summed over B) 

prop.table(mytable.networth, 1) # row percentages 

```

```{r}
#Ethnicity vs. Party 
qplot(factor(ethnicity), data=vf, geom="bar", fill=factor(party))
mytable.ethnicity <- table(vf$party, vf$ethnicity) # A is row, B is columns 
mytable.ethnicity

margin.table(mytable.ethnicity, 1) # A frequencies (summed over B) 

prop.table(mytable.ethnicity, 1) # row percentages 

#Digging Deeper in Hispanic Population 
# Hispanic Only dataset
df <- vf[vf$ethnicity == "Hispanic",] 
mytable.hispanic <- table(df$party, df$ethnicity)
mytable.hispanic
prop.table(mytable.hispanic, 2)
```


```{r}
#Marital Status vs. Party 
qplot(factor(maritalstatus), data=vf, geom="bar", fill=factor(party))
mytable.marital <- table(vf$party, vf$maritalstatus) # A is row, B is columns 
mytable.marital

margin.table(mytable.marital, 1) # A frequencies (summed over B) 

prop.table(mytable.marital, 1) # row percentages 
```

```{r}
# Homeowner status vs. Party
qplot(factor(home_owner_or_renter), data=vf, geom="bar", fill=factor(party))
mytable.home <- table(vf$party, vf$home_owner_or_renter) # A is row, B is columns 
mytable.home

margin.table(mytable.home, 1) # A frequencies (summed over B) 

prop.table(mytable.home, 1) # row percentages 
```

```{r}
# Congressional District vs. Party
qplot(factor(cd), data=vf, geom="bar", fill=factor(party))
mytable.cd <- table(vf$party, vf$cd) # A is row, B is columns 
mytable.cd

margin.table(mytable.cd, 1) # A frequencies (summed over B) 

prop.table(mytable.cd, 2) # column percentages 
```

**Non-partisan voters**
We see that the average non-partisan voter is 

* European 

* Lives in Single Family Dwelling Unit

* Almost equally scattered in all 4 districts 

* More likely married than non-traditional 

* Making between $100,000-249,999
```{r}
#Digging Deeper in Non-partisan populations
# creating dataset with only non-part. voters 
df <- vf[vf$party == "Non-Partisan",] 
mytable.np <- table(df$party, df$ethnicity)
#Proportions across ethnicites 
prop.table(mytable.np, 1)
#Proportions across dwelling 
mytable.np1 <- table(df$party, df$dwellingtype) 
prop.table(mytable.np1, 1) 
#Proportions across Congressional District 
mytable.np2 <- table(df$party, df$cd) 
prop.table(mytable.np2, 1) 
#Proportions across marital status
mytable.np2 <- table(df$party, df$maritalstatus) 
prop.table(mytable.np2, 1) 
#Proportions across networth
mytable.np2 <- table(df$party, df$net_worth) 
prop.table(mytable.np2, 1) 
```


